Sex, education, and mental health: Analyzing American census data through statistical methods, 2018 to 2021

### 2024-10-25 paper/paper.qmd

---
title: "Gender and Mental Depression: Analyzing Census Data Through Statistical Methods"
author: 
  - Christina Nguyen
  - Marcus Barnes
  - Nibras Ar Rakib
thanks: "Code and data are available at: [https://github.com/Nguyen-Ar-Rakib-Barnes/downturns-household-insurance-and-poverty](https://github.com/Nguyen-Ar-Rakib-Barnes/downturns-household-insurance-and-poverty)."
date: today
date-format: long
abstract: "Mental health posed a significant global health threat before the pandemic, and recent studies have shown that it has become even more prevalent. Our study developed a method to identify severe psychological depression using survey data that can be used as an early alert system to identify nonspecific psychological distress."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

combined <- bind_rows(count_2018, count_2021)
max_count <-  max(combined$Count) + 1000
ggplot(combined, aes(x = sex, y = Count, fill = Year)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.6) +
  scale_y_continuous(limits=c(0,max_count)) + 
    geom_text(aes(label = Count), 
            position = position_dodge(width = 0.6),
            vjust = -0.3,
            size = 5) +
  labs(
    title = "Gender Distribution: 2018 vs 2021",
    x = "Gender",
    y = "Count"
  ) +
  theme_light() +
  theme(
    plot.title = element_text(size = 18, face = "bold"),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16)
  )
```

>>>>

---
title: "Sex, education, and mental health: Analyzing American census data through statistical methods, 2018 to 2021"
subtitle: "A pre- and post-pandemic study with IPUM's MEPs data"
author: 
  - Christina Nguyen
  - Marcus Barnes
  - Nibras Ar Rakib
thanks: "Code and data are available at: [https://github.com/Nguyen-Ar-Rakib-Barnes/downturns-household-insurance-and-poverty](https://github.com/Nguyen-Ar-Rakib-Barnes/downturns-household-insurance-and-poverty)."
date: today
date-format: long
abstract: "Mental health posed a significant global health threat before the pandemic, and recent studies have shown that it has become even more prevalent. Our study developed a method to identify severe psychological depression using survey data that can be used as an early alert system to identify nonspecific psychological distress."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

combined <- bind_rows(count_2018, count_2021)
max_count <-  max(combined$Count) + 1000
ggplot(combined, aes(x = sex, y = Count, fill = Year)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.6) +
  scale_y_continuous(limits=c(0,max_count)) + 
    geom_text(aes(label = Count), 
            position = position_dodge(width = 0.6),
            vjust = -0.3,
            size = 5) +
  labs(
    title = "Gender Distribution: 2018 vs 2021",
    x = "Gender",
    y = "Count"
  ) +
  theme(
    plot.title = element_text(size = 18, face = "bold"),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16)
  )
```
